---
title: "Create datasets for supplemental analyses"
author: "Roland Knapp"
date: "`r Sys.Date()`"
output: html_notebook
---

## Load packages

```{r load-packages}
library(RPostgreSQL)
library(dplyr) 
library(readr)
library(tibble)
library(tidyr)
library(lubridate)
```

-   Users without credentials to access database can utilize `subadult_survival.csv` and `tadpole_survival.csv` saved below.

## DATASET 1: Bd load before versus after translocation

-   Loads at time of translocation versus from CMR years 0 and 1. Exclude years \> 1. Only frogs that were recaptured at least once following translocation are included.

### Read in raw translocation and capture datasets, bind them

```{r}
translocation <- read_csv(here::here("data", "raw", "frog_translocation.csv")) %>% 
  relocate(visit_date = release_date, .before = type) %>% 
  rename(survey_type = type) %>% 
  mutate(pit_tag_ref = replace(pit_tag_ref, pit_tag_ref == "900067000117396" & site_id == 70641, "111111111111111")) %>% # remove incorrect pit_tag_ref identified in translocation_survival_createdata.Rmd
  filter(pit_tag_ref != "111111111111111")

capture <- read_csv(here::here("data", "raw", "frog_captures.csv")) %>% 
  select(-capture_life_stage, -capture_animal_state, -sex)

# select only translocated frogs (exclude naturally-recruited frogs and translocated frogs that were never recaptured after translocation)
capture <- translocation %>% 
  select(pit_tag_ref) %>% 
  inner_join(capture, by = c("pit_tag_ref")) %>% 
  relocate(pit_tag_ref, .after = last_col())

frogs_all <- bind_rows(translocation, capture)
```

### Connect to PostgreSQL database `amphibians`

```{r}
source("db_connect.R")
```

### Retrieve Bd data for translocated frogs

```{sql retrieve-translocation-bd-data, connection=con, output.var="translocate_bd"}
select
  release_date as visit_date,
  pit_tag_ref,
  swab_id,
  bd_load
from relocate
inner join relocate_frog on relocate.id = relocate_id
inner join bd_load on swab_id = sample_id  
where release_siteid1 in (70134, 70279, 70370, 70413, 70414, 70449, 70505, 70550, 70556, 70619, 70628, 70641, 74976) and
  (replicate = 1 or replicate is null)
```

### Retrieve Bd data for captured frogs

```{sql retrieve-capture-bd-data, connection=con, output.var="capture_bd"}
select
	visit_date,
	pit_tag_ref, 
	swab_id,
	bd_load
from visit
inner join survey on visit.id = visit_id
inner join capture_survey on survey.id = survey_id
inner join bd_load on swab_id = sample_id  
where site_id in (70134, 70688, 70370, 70114, 70175, 70279, 70413, 71570, 71968, 72008, 72093, 72264, 72390, 72442, 72694, 70414, 70449, 70034, 70505, 72092, 70550, 70556, 70619, 70628, 70641, 74976, 72973) and 
	survey_type = 'cmr' and
	capture_animal_state != 'dead' and
	(replicate = 1 or replicate is null)
```

### Disconnect from database

```{r}
source("db_disconnect.R")
```

### Combine Bd data from translocate and capture datasets

```{r}
bd_add <- bind_rows(translocate_bd, capture_bd)
```

### Add Bd data to frog data, restructure dataset

```{r}
frogsbd <- frogs_all %>% 
  inner_join(bd_add, by = c("visit_date", "pit_tag_ref")) %>% 
  mutate(period = if_else(survey_type == "cmr", "after", "before")) %>% 
  relocate(period, .after = survey_type) %>% 
  arrange(site_id, visit_date)
```

### Add years since translocation

```{r}
frogsbd <- translocation %>% 
  mutate(year_translocate = year(visit_date)) %>% 
  select(year_translocate, pit_tag_ref) %>% 
  inner_join(frogsbd, by = c("pit_tag_ref")) %>% 
  relocate(year_translocate, .after = visit_date) %>% 
  relocate(pit_tag_ref, .after = period) %>% 
  mutate(year_capture = year(visit_date), .after = year_translocate) %>% 
  mutate(interval = (year_capture - year_translocate), .after = year_capture) %>% 
  arrange(site_id, visit_date, pit_tag_ref)
```

### Update site ids to standardize usage

```{r}
frogsbd <- frogsbd %>% 
  mutate(site_id1 = site_id, .after = site_id,
         site_id1 = replace(site_id1, site_id == 70034, 70449),
         site_id1 = replace(site_id1, site_id == 70175, 70413),
         site_id1 = replace(site_id1, site_id == 70279, 70413),
         site_id1 = replace(site_id1, site_id == 72093, 70413),
         site_id1 = replace(site_id1, site_id == 72442, 70413),
         site_id1 = replace(site_id1, site_id == 70688, 70134),
         site_id1 = replace(site_id1, site_id == 72092, 70505),
         site_id1 = replace(site_id1, site_id == 72973, 74976))
```

### Add translocation_id

```{r}
frogsbd <- frogsbd %>% 
  distinct(site_id1, year_translocate) %>% 
  arrange(site_id1, year_translocate) %>% 
  group_by(site_id1) %>% 
  mutate(transno = seq_len(n())) %>% 
  ungroup(site_id1) %>% 
  unite(translocation_id, c("site_id1", "transno"), remove = FALSE) %>% 
  select(-transno) %>% 
  inner_join(frogsbd, by = c("site_id1", "year_translocate")) %>% 
  select(site_id, site_id1, translocation_id, visit_date, survey_type, period, year_translocate, year_capture, interval, pit_tag_ref, swab_id, bd_load) %>% 
  arrange(site_id1, visit_date)
```

### Save frog-Bd dataset

```{r}
frogsbd %>% write_csv(here::here("data", "clean", "bd_beforeafter_translocation.csv"))
```

### Description of dataset columns

-   site_id: Unique 5-digit site identifier.
-   site_id1: Assigned 5-digit site identifier for a collection of adjacent sites that are utilized by a frog population.
-   translocation_id: Unique identifier for each translocation.
-   visit_date: Date on which translocation or survey was conducted.
-   survey_type: "translocation" or "cmr".
-   period: "before" or "after" translocation.
-   year_translocate: Year in which translocation was conducted.
-   year_capture: Year in which frog was captured.
-   interval: Number of years between translocation and frog capture.
-   pit_tag_ref: Unique PIT tag identifier.
-   swab_id: Unique identifier for skin swab sample.
-   bd_load: Bd infection intensity determined from skin swab samples, units are number of ITS1 copies \* 80.

## DATASETS 2: Create datasets for use in estimating selected parameters for viability analysis

### Estimate subadult survival
-   Used 70550 because of its long and relatively intensive survey history.

#### Connect to database

```{r}
source("db_connect.R")
```

#### Retrieve subadult count data

```{sql, connection=con, output.var="surv"}
select
  site_id,
  visit_date,
  species,
  visual_life_stage,
  visual_animal_state,
  count
from visit
inner join survey on visit_id = visit.id
inner join visual_survey on survey_id = survey.id  
where site_id in (70550) and
  species = 'ramu' and
  visual_life_stage = 'subadult' and
  visual_animal_state = 'alive'
order by site_id, visit_date
```

#### Disconnect from database

```{r}
source("db_disconnect.R")
```

#### Save raw dataset

```{r}
surv %>% write_csv(here::here("data", "raw", "subadult_survival.csv"))
```

#### Calculate maximum annual counts

```{r}
surv <- surv %>% 
  group_by(visit_date) %>% 
  summarize(count_sum = sum(count)) %>% 
  ungroup() %>% 
  mutate(year = year(visit_date)) %>% 
  group_by(year) %>% 
  summarize(count_max = max(count_sum))
```

#### Extract number of new recruits from mrmr model

```{r}
model <- read_rds(here::here("data", "clean", "70550_model.rds"))

get_numeric_indices <- function(string) {
  idx_mat <- stringr::str_extract_all(string, "[0-9]+", simplify = TRUE)
  colnames(idx_mat) <- paste0("index_", seq_len(ncol(idx_mat)))
  tibble::as_tibble(idx_mat) %>%
    dplyr::mutate_all(readr::parse_integer)
}

survey_prim_periods <- model$data$surveys %>%
  group_by(.data$primary_period) %>%
  filter(.data$secondary_period == min(.data$secondary_period)) %>% 
  ungroup()

p <- model$m_fit$draws("B", format = "draws_df") %>%
  tidyr::pivot_longer(tidyselect::starts_with("B")) %>%
  suppressWarnings() %>%
  mutate(
    get_numeric_indices(.data$name),
    primary_period = .data$index_1 + 1
  ) %>%
  group_by(.data$primary_period) %>%
  summarize(lo = quantile(.data$value, .025),
            med = median(.data$value),
            hi = quantile(.data$value, .975),
            .groups = "drop") %>%
  left_join(survey_prim_periods, by = "primary_period")

recruit <- p %>% 
  group_by(year) %>% 
  summarize(recruit_sum = sum(med))
```

#### Create final dataset

```{r}
subadult_survival <- surv %>% 
  inner_join(recruit, by = c("year"))

subadult_survival %>% write_csv(here::here("data", "clean", "subadult_survival.csv"))
```

#### Calculate subadult survival

-   Assumes that p(detection) for subadults is 0.25.

```{r}
subadult_sum <- subadult_survival %>% 
  filter(between(year, 2008, 2020)) %>% 
  summarize(sum(count_max))

recruit_sum <- subadult_survival %>% 
  filter(between(year, 2009, 2021)) %>% 
  summarize(sum(recruit_sum))

surv_sub <- recruit_sum / (subadult_sum * 4) 
```

### Estimate 1st-year tadpole survival

-   Used 74976 due to its shallow habitats that allow accurate tadpole counts.

-   1st-year tadpole counts are from Knapp field notebooks - VES counts in database do not distinguish between tadpole age-classes. 

```{r}
surv_tad <- tibble(site_id = 74976, year = 2019:2021, count = c(250, 300, 305))
```

#### Extract adult abundance from mrmr model

-   Used minimum survey date per year to estimate number of females during breeding season.

```{r}
model <- read_rds(here::here("data", "clean", "74976_model.rds"))

get_numeric_indices <- function(string) {
  idx_mat <- stringr::str_extract_all(string, "[0-9]+", simplify = TRUE)
  colnames(idx_mat) <- paste0("index_", seq_len(ncol(idx_mat)))
  tibble::as_tibble(idx_mat) %>%
    dplyr::mutate_all(readr::parse_integer)
}

survey_prim_periods <- model$data$surveys %>%
  group_by(.data$primary_period) %>%
  filter(.data$secondary_period == min(.data$secondary_period)) %>% 
  ungroup()

p <- model$m_fit$draws("N", format = "draws_df") %>%
        tidyr::pivot_longer(tidyselect::starts_with("N")) %>%
        suppressWarnings() %>%
        mutate(
          get_numeric_indices(.data$name),
          primary_period = .data$index_1 + 1
        ) %>%
        group_by(.data$primary_period) %>%
        summarize(lo = quantile(.data$value, .025),
                  med = median(.data$value),
                  hi = quantile(.data$value, .975),
                  .groups = "drop") %>%
        left_join(survey_prim_periods, by = "primary_period")

abund_ad <- p %>% 
  group_by(year) %>% 
  summarize(survey_date = min(survey_date)) %>%
  select(survey_date) %>% 
  inner_join(p, by = c("survey_date")) %>% 
  select(year, med)
```

#### Create final dataset

```{r}
surv_tad <- surv_tad %>% 
  inner_join(abund_ad, by = "year") %>% 
  rename(adults = med)

surv_tad %>% write_csv(here::here("data", "clean", "tadpole_survival.csv"))
```

#### Calculate tadpole survival

Assumptions:

-   p(detection) for 1st-year tadpoles is 0.5.

-   \# females is half of total adult abundance

-   female fecundity = 100 eggs

```{r}
surv_tad <- surv_tad %>% 
  mutate(surv_1styr = (count * 2) / ((adults / 2) * 100)) %>% 
  summarize(mean(surv_1styr))
```

## DATASETS 3: Create datasets for maps

### Dataset for map of sample sites for exome capture analysis

Users without the necessary credentials to access the database can access the final dataset at data/maps/exome_points.csv.

```{r}
exome_sites <- read_csv(here::here("data", "maps", "exome_sample_sites.csv"))
n <- nrow(exome_sites)

source("db_connect.R")
exome_points <- dbReadTable(con, c("site")) %>% 
  as_tibble() %>% 
  rename(site_id = id) %>% 
  select(site_id, utme, utmn) %>% 
  inner_join(exome_sites, by = "site_id") %>% 
  add_column(x = sample(-2000:2000, size = n), # Add x-y random noise to obscure exact site locations
             y = sample(-2000:2000, size = n)) %>% 
  mutate(utme_noise = utme + x,
         utmn_noise = utmn + y) %>% 
  select(-utme, -utmn, -x, -y)
source("db_disconnect.R")

write_csv(exome_points, here::here("data", "maps", "exome_points.csv"))
```

## Dataset for map of Yosemite translocation & donor sites

Users without the necessary credentials to access the database can access the final dataset at data/maps/translocation_points.csv.

### Retrieve data & create points layer

```{r}
donor_sites <- read_csv(here::here("data", "clean", "frog_translocation_final.csv")) %>% 
  distinct(donor) %>% 
  mutate(site_id = donor, .before = donor) %>% 
  rename(donor_id = donor) %>% 
  add_column(type = "donor")

translocation_sites <- read_csv(here::here("data", "clean", "frog_translocation_final.csv")) %>% 
  distinct(site_id, donor) %>% 
  rename(donor_id = donor) %>% 
  add_column(type = "translocation") %>% 
  bind_rows(donor_sites)

n <- nrow(translocation_sites)

rm(donor_sites)

source("db_connect.R")
translocation_points <- dbReadTable(con, c("site")) %>% 
  as_tibble() %>% 
  rename(site_id = id) %>% 
  select(site_id, utme, utmn) %>% 
  inner_join(translocation_sites, by = "site_id") %>% 
  add_column(x = sample(-2000:2000, size = n), # Add x-y random noise to obscure exact site locations
             y = sample(-2000:2000, size = n)) %>% 
  mutate(utme_noise = utme + x,
         utmn_noise = utmn + y) %>% 
  select(-utme, -utmn, -x, -y)
source("db_disconnect.R")

write_csv(translocation_points, here::here("data", "maps", "translocation_points.csv"))
```
