---
title: "Predictors of frog survival following translocations - supplemental analyses"
author: "Roland Knapp"
output: html_notebook
---

The following are supplemental analyses, some related to analyses in `translocation_survival_analysis.Rmd` and some unrelated.

## Load packages

```{r}
library(RPostgreSQL)
library(tidyverse)
library(lubridate)
library(brms)
library(rstanarm)
library(bayesplot)
library(tidybayes)
library(loo)
library(broom.mixed)
library(janitor)

source("classification_summary.R")
```

## ANALYSIS 1: Analysis of changes in Bd load following translocation

-   Compare loads at time of translocation to those from CMR years 0 and 1. Exclude years \> 1. Only frogs that were recaptured at least once following translocation are included.

### Create dataset

### Read in raw translocation and capture datasets, bind them

```{r}
translocation <- read_csv(here::here("data", "raw", "frog_translocation.csv")) %>% 
  relocate(visit_date = release_date, .before = type) %>% 
  rename(survey_type = type) %>% 
  mutate(pit_tag_ref = replace(pit_tag_ref, pit_tag_ref == "900067000117396" & site_id == 70641, "111111111111111")) %>% # remove incorrect pit_tag_ref identified in translocation_survival_createdata.Rmd
  filter(pit_tag_ref != "111111111111111")

capture <- read_csv(here::here("data", "raw", "frog_captures.csv")) %>% 
  select(-capture_life_stage, -capture_animal_state, -sex)

# select only translocated frogs (exclude naturally-recruited frogs and translocated frogs that were never recaptured after translocation)
capture <- translocation %>% 
  select(pit_tag_ref) %>% 
  inner_join(capture, by = c("pit_tag_ref")) %>% 
  relocate(pit_tag_ref, .after = last_col())

frogs_all <- bind_rows(translocation, capture)
```

### Connect to PostgreSQL database `amphibians`

```{r}
con = dbConnect(dbDriver("PostgreSQL"),
                user = rstudioapi::askForPassword("user name"),
                password = rstudioapi::askForPassword("password"),
                host = "frogdb.eri.ucsb.edu",
                port = 5432,
                dbname = "amphibians")

```

### Retrieve Bd data for translocated frogs from database

```{sql retrieve-translocation-bd-data, connection=con, output.var="translocate_bd"}
select
  release_date as visit_date,
  pit_tag_ref,
  swab_id,
  bd_load
from relocate
inner join relocate_frog on relocate.id = relocate_id
inner join bd_load on swab_id = sample_id  
where release_siteid1 in (70134, 70279, 70370, 70413, 70414, 70449, 70505, 70550, 70556, 70619, 70628, 70641, 74976) and
  (replicate = 1 or replicate is null)
```

### Retrieve Bd data for captured frogs from database

```{sql retrieve-capture-bd-data, connection=con, output.var="capture_bd"}
select
	visit_date,
	pit_tag_ref, 
	swab_id,
	bd_load
from visit
inner join survey on visit.id = visit_id
inner join capture_survey on survey.id = survey_id
inner join bd_load on swab_id = sample_id  
where site_id in (70134, 70688, 70370, 70114, 70175, 70279, 70413, 71570, 71968, 72008, 72093, 72264, 72390, 72442, 72694, 70414, 70449, 70034, 70505, 72092, 70550, 70556, 70619, 70628, 70641, 74976, 72973) and 
	survey_type = 'cmr' and
	capture_animal_state != 'dead' and
	(replicate = 1 or replicate is null)
```

### Disconnect from database

```{r}
dbDisconnect(con) 
rm("con")
```

### Combine Bd data from translocate and capture datasets

```{r}
bd_add <- bind_rows(translocate_bd, capture_bd)
```

### Add Bd data to frog data, restructure dataset

```{r}
frogsbd <- frogs_all %>% 
  inner_join(bd_add, by = c("visit_date", "pit_tag_ref")) %>% 
  mutate(period = if_else(survey_type == "cmr", "after", "before")) %>% # use 0/1 instead?
  relocate(period, .after = survey_type) %>% 
  arrange(site_id, visit_date)
```

### Add years since translocation

```{r}
frogsbd <- translocation %>% 
  mutate(year_translocate = year(visit_date)) %>% 
  select(year_translocate, pit_tag_ref) %>% 
  inner_join(frogsbd, by = c("pit_tag_ref")) %>% 
  relocate(year_translocate, .after = visit_date) %>% 
  relocate(pit_tag_ref, .after = period) %>% 
  mutate(year_capture = year(visit_date), .after = year_translocate) %>% 
  mutate(interval = (year_capture - year_translocate), .after = year_capture) %>% 
  arrange(site_id, visit_date, pit_tag_ref)
```

### Update site ids to standardize usage

```{r}
frogsbd <- frogsbd %>% 
  mutate(site_id1 = site_id, .after = site_id,
         site_id1 = replace(site_id1, site_id == 70034, 70449),
         site_id1 = replace(site_id1, site_id == 70175, 70413),
         site_id1 = replace(site_id1, site_id == 70279, 70413),
         site_id1 = replace(site_id1, site_id == 72093, 70413),
         site_id1 = replace(site_id1, site_id == 72442, 70413),
         site_id1 = replace(site_id1, site_id == 70688, 70134),
         site_id1 = replace(site_id1, site_id == 72092, 70505),
         site_id1 = replace(site_id1, site_id == 72973, 74976))
```

### Add translocation_id

```{r}
frogsbd <- frogsbd %>% 
  distinct(site_id1, year_translocate) %>% 
  arrange(site_id1, year_translocate) %>% 
  group_by(site_id1) %>% 
  mutate(transno = seq_len(n())) %>% 
  ungroup(site_id1) %>% 
  unite(translocation_id, c("site_id1", "transno"), remove = FALSE) %>% 
  select(-transno) %>% 
  inner_join(frogsbd, by = c("site_id1", "year_translocate")) %>% 
  select(site_id, site_id1, translocation_id, visit_date, survey_type, period, year_translocate, year_capture, interval, pit_tag_ref, swab_id, bd_load) %>% 
  arrange(site_id1, visit_date)
```

### Save frog-Bd dataset

```{r}
frogsbd %>% write_csv(here::here("data", "clean", "bd_beforeafter_translocation.csv"))
```

### Description of dataset columns

-   site_id: Unique 5-digit site identifier.
-   site_id1: Assigned 5-digit site identifier for a collection of adjacent sites that are utilized by a frog population.
-   translocation_id: Unique identifier for each translocation.
-   visit_date: Date on which translocation or survey was conducted.
-   survey_type: "translocation" or "cmr".
-   period: "before" or "after" translocation.
-   year_translocate: Year in which translocation was conducted.
-   year_capture: Year in which frog was captured.
-   interval: Number of years between translocation and frog capture.
-   pit_tag_ref: Unique PIT tag identifier.
-   swab_id: Unique identifier for skin swab sample.
-   bd_load: Bd infection intensity determined from skin swab samples, units are number of ITS1 copies \* 80.

### Analyze dataset

-   Model with pit_tag_ref grouping variable is the model of primary interest, roughly analogous to paired t-test. Included other models to allow a better understanding of the data.

#### Read in data

```{r}
d1 <- read_csv(here::here("data", "clean", "bd_beforeafter_translocation.csv")) 
```

#### Data exploration

```{r}
d1 %>% 
  mutate(period = factor(period, levels = c("before", "after"))) %>% 
  ggplot(aes(x = period, y = log10(bd_load + 1))) +
          geom_boxplot()

d1 %>% 
  mutate(period = factor(period, levels = c("before", "after"))) %>% 
  ggplot(aes(x = period, y = log10(bd_load + 1))) +
    geom_boxplot() + 
    labs(x = "Period", y = "Bd load (log10(bd_load + 1))") +
    facet_grid(. ~ site_id1) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5),
          strip.text.x = element_text(size = 8),
          panel.grid = element_blank())
ggsave(here::here("doc", "manuscript", "figures", "bdload_beforeafter.png"), width = 6, height = 3)
```

-   In most cases, relatively small increases in load following translocation.

#### Prepare variables for analysis

```{r}
d1 <- d1 %>% 
  mutate(after = if_else(period == "after", 1, 0),
         bd_load = as.integer(round(bd_load)), 
         after_c = after - mean(after)) 
```

#### Check for null values

```{r check-null}
d1 %>% summarize(across(everything(), ~ sum(is.na(.))))
```

#### Include only CMR years 0 and 1

```{r}
d2 <- d1 %>% 
  filter(interval < 2)
```

#### Plot distribution of bd_load response variable

```{r}
d2 %>% 
  ggplot(aes(x = bd_load)) +
    geom_histogram()

d2 %>% 
  ggplot(aes(sample = bd_load)) +
    geom_qq() +
    geom_qq_line()
```

-   Data are strongly left-skewed, possibly zero-inflated. Model using negative binomial family.

### Model without group-level effects

#### Model specification

```{r}
m1a <- stan_glm(
  bd_load ~ after_c,
  data = d2,
  family = neg_binomial_2(),
  chains = 4, 
  iter = 5000*2,
  cores = 4,
  seed = 84735)
```

#### Priors and chain diagnostics

```{r}
prior_summary(m1a)

mcmc_trace(m1a, size = 0.1)
mcmc_dens_overlay(m1a)

summary(m1a)
```

#### Posterior predictive check

```{r}
pp_check(m1a, plotfun = "stat") +
    xlab("Bd load")
```

-   Posterior predictive distribution implied by regression model closely matches original data (based on mean Bd load).

#### Calculate loo

```{r}
loo_m1a <- loo(m1a, cores = 4, save_psis = TRUE)
print(loo_m1a)
```

#### Posterior analysis

```{r}
tidy(m1a, conf.int = TRUE, conf.level = 0.90)

mcmc_areas(m1a, area_method = "equal height", prob = 0.90) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors, no group-level effects",
          "Posterior distributions with medians & 90% uncertainty intervals")

# mean(bayes_R2(m1a))   # bayes_R2 not available for negative binomials models in stanreg. 
```

### Model with 1 grouping variable: pit_tag_ref

#### Model specification

```{r}
m1b <- stan_glmer(
  bd_load ~ after_c + (1 | pit_tag_ref),
  data = d2,
  family = neg_binomial_2(),
  chains = 4, 
  iter = 5000*2,
  cores = 4,
  seed = 84735)
```

#### List model variable names

```{r}
get_variables(m1b)
```

#### Priors and chain diagnostics

```{r}
prior_summary(m1b)

b1 <- c("(Intercept)", "after_c")

mcmc_trace(m1b, size = 0.1, pars = b1)
mcmc_dens_overlay(m1b, pars = b1)

summary(m1b)
```

#### Posterior predictive check

```{r}
pp_check(m1b, plotfun = "stat") +
    xlab("Bd load")
```

-   Posterior predictive distribution implied by regression model shows some mismatch between actual data and model results (based on mean Bd load).

#### Calculate loo

```{r}
loo_m1b <- loo(m1b, cores = 4, save_psis = TRUE)
print(loo_m1b)
```

#### Posterior analysis

```{r}
tidy(m1b, conf.int = TRUE, conf.level = 0.90)

mcmc_areas(m1b, area_method = "equal height", prob = 0.90, pars = b1) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with pit_tag_ref as grouping variable",
          "Posterior distributions with medians & 90% uncertainty intervals")

# mean(bayes_R2(m1b))
```

### Model with 2 grouping variables: pit_tag_ref nested within translocation_id

#### Model specification

```{r}
m1c <- stan_glmer(
  bd_load ~ after_c + (1 | pit_tag_ref) + (1 | translocation_id),
  data = d2,
  family = neg_binomial_2(),
  chains = 4, 
  iter = 5000*2,
  cores = 4,
  seed = 84735)
```

#### Priors and chain diagnostics

```{r}
prior_summary(m1c)

b1 <- c("(Intercept)", "after_c")

mcmc_trace(m1c, size = 0.1, pars = b1)
mcmc_dens_overlay(m1c, pars = b1)

summary(m1c)
```

#### Posterior predictive check

```{r}
pp_check(m1c, plotfun = "stat") +
    xlab("Bd load")
```

-   Posterior predictive distribution implied by regression model closely matches original data (based on mean Bd load).

#### Calculate loo

```{r}
loo_m1c <- loo(m1c, cores = 4, save_psis = TRUE)
print(loo_m1c)
```

#### Posterior analysis

```{r}
tidy(m1c, conf.int = TRUE, conf.level = 0.90)

mcmc_areas(m1c, area_method = "equal height", prob = 0.90, pars = b1) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with pit_tag_ref and translocation_id as grouping variables",
          "Posterior distributions with medians & 90% uncertainty intervals")

# mean(bayes_R2(m1c))
```

### Model with 2 grouping variables: pit_tag_ref nested within site_id1

#### Model specification

```{r}
m1d <- stan_glmer(
  bd_load ~ after_c + (1 | pit_tag_ref) + (1 | site_id1),
  data = d2,
  family = neg_binomial_2(),
  chains = 4, 
  iter = 5000*2,
  cores = 4,
  seed = 84735)
```

#### Priors and chain diagnostics

```{r}
prior_summary(m1d)

b1 <- c("(Intercept)", "after_c")

mcmc_trace(m1d, size = 0.1, pars = b1)
mcmc_dens_overlay(m1d, pars = b1)

summary(m1d)
```

#### Posterior predictive check

```{r}
pp_check(m1d, plotfun = "stat") +
    xlab("Bd load")
```

-   Posterior predictive distribution implied by regression model closely matches original data (based on mean Bd load).

#### Calculate loo

```{r}
loo_m1d <- loo(m1d, cores = 4, save_psis = TRUE)
print(loo_m1d)
```

#### Posterior analysis

```{r}
tidy(m1d, conf.int = TRUE, conf.level = 0.90)

mcmc_areas(m1d, area_method = "equal height", prob = 0.90, pars = b1) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with pit_tag_ref and site_id1 as grouping variables",
          "Posterior distributions with medians & 90% uncertainty intervals")

# mean(bayes_R2(m1d))
```

### Compare 4 models with different grouping structures

```{r}
loo_compare(loo_m1a, loo_m1b, loo_m1c, loo_m1d)
```

-   All models are equivalent, so pairing before-after by pit_tag_ref is unimportant.

## ANALYSIS 2: Evaluate effect of frog size independent of donor population

-   Model m1d from `translocation_survival_analysis.Rmd` but using data subset where donor = 72996 only.

### Read in frog translocation dataset

-   arm package used to rescale continuous variables (substract mean, divide by 2 SD), manual centering of binary variables.

```{r read-translocation-data}
d1 <- read_csv(here::here("data", "clean", "frog_translocation_final.csv")) %>% 
  drop_na(bd_load) %>%  # drop 12 records where bd_load is NA (9 are from 70413)
  mutate(
    first = if_else(order == 1, 1, 0),
    surv = if_else(survival < 0.5, 0, 1),
    male = if_else(sex == "m", 1, 0),
    elev_lo = if_else(elevation < 3020, 1, 0), # using interval defined by cut_interval, n = 2
    elev_z = arm::rescale(elevation),
    snowt_z = arm::rescale(snow_t),
    snowt1_z = arm::rescale(snow_t1),
    day_z = arm::rescale(day),
    length_z = arm::rescale(length),
    bdload_l = log10(bd_load + 1),
    bdload_z = arm::rescale(bdload_l),
    shore_c = shore - mean(shore),
    male_c = male - mean(male),
    elevlo_c = elev_lo - mean(elev_lo),
    first_c = first - mean(first),
    across(c(shore_c, male_c, elevlo_c, first_c), round, 3), 
    across(c(site_id, shore_c, first_c, donor, male_c, elevlo_c, year), as.factor),
    surv = as.integer(surv))

# Add a descriptive translocation_id
d1 <- d1 %>% 
  distinct(site_id, year) %>% 
  arrange(site_id, year) %>% 
  group_by(site_id) %>% 
  mutate(transno = seq_len(n())) %>% 
  ungroup(site_id) %>% 
  unite(translocation_id, c("site_id", "transno"), remove = FALSE) %>% 
  select(-transno) %>% 
  mutate(translocation_id = as.factor(translocation_id)) %>% 
  inner_join(d1, by = c("site_id", "year")) %>% 
  relocate(translocation_id, .after = site_id)
```

### Model specification

```{r}
d1_72996 <- d1 %>% 
  filter(donor == "72996" & !is.na(bd_load))

m1d_72996 <- stan_glmer(
  surv ~ length_z + snowt_z + snowt1_z + day_z + bdload_z + elev_z + first_c + male_c + (1 | translocation_id),
  data = d1_72996,
  family = binomial,
  chains = 4, 
  iter = 5000*2, 
  cores = 4,
  seed = 84735)
```

### Priors & chain diagnostics

```{r}
prior_summary(m1d_72996)

b1_72996 <- c("(Intercept)", "length_z", "snowt_z", "snowt1_z", "day_z", "bdload_z", "elev_z", "first_c0.353", "male_c0.548")

mcmc_trace(m1d_72996, size = 0.1, pars = b1_72996)
mcmc_dens_overlay(m1d_72996, pars = b1_72996)

summary(m1d_72996)
```

### Posterior predictive check

```{r}
pp_check(m1d_72996, plotfun = "stat") +
    xlab("Frog survival rate")
```

### Posterior analysis

```{r}
tidy(m1d_72996, conf.int = TRUE, conf.level = 0.90)

mcmc_areas(m1d_72996, area_method = "equal height", prob = 0.90, pars = b1_72996) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors & translocation_id as grouping variable",
          "Posterior distributions with medians & 90% uncertainty intervals")

mean(bayes_R2(m1d_72996))

# Frog survival by translocation_id
mcmc_areas_ridges(m1d_72996, regex_pars = "translocation_id") +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Frog survival for each translocation_id")
```

-   Using dataset containing only frogs translocated from 72996 had relatively little effect on the predictors of frog survival. Specific to the objective of this analysis, frog size remained unimportant as a predictor of survival. However, effect of Bd load became marginally unimportant instead of unimportant (i.e., uncertainty interval includes 0 but barely).

## ANALYSIS 3: Assess effect on m1d of including all records, including where bd_load = NA

-   Additional analysis of model m1d from `translocation_survival_analysis.Rmd`.

### Read in frog translocation dataset

-   arm package used to rescale continuous variables (substract mean, divide by 2 SD), manual centering of binary variables.

```{r read-translocation-data-including-na}
d1_allna <- read_csv(here::here("data", "clean", "frog_translocation_final.csv")) %>% 
#  drop_na(bd_load) %>%  # drop 12 records where bd_load is NA (9 are from 70413)
  mutate(
    first = if_else(order == 1, 1, 0),
    surv = if_else(survival < 0.5, 0, 1),
    male = if_else(sex == "m", 1, 0),
    elev_lo = if_else(elevation < 3020, 1, 0), # using interval defined by cut_interval, n = 2
    elev_z = arm::rescale(elevation),
    snowt_z = arm::rescale(snow_t),
    snowt1_z = arm::rescale(snow_t1),
    day_z = arm::rescale(day),
    length_z = arm::rescale(length),
    bdload_l = log10(bd_load + 1),
    bdload_z = arm::rescale(bdload_l),
    shore_c = shore - mean(shore),
    male_c = male - mean(male),
    elevlo_c = elev_lo - mean(elev_lo),
    first_c = first - mean(first),
    across(c(shore_c, male_c, elevlo_c, first_c), round, 3), 
    across(c(site_id, shore_c, first_c, donor, male_c, elevlo_c, year), as.factor),
    surv = as.integer(surv))

# Add a descriptive translocation_id
d1_allna <- d1_allna %>% 
  distinct(site_id, year) %>% 
  arrange(site_id, year) %>% 
  group_by(site_id) %>% 
  mutate(transno = seq_len(n())) %>% 
  ungroup(site_id) %>% 
  unite(translocation_id, c("site_id", "transno"), remove = FALSE) %>% 
  select(-transno) %>% 
  mutate(translocation_id = as.factor(translocation_id)) %>% 
  inner_join(d1_allna, by = c("site_id", "year")) %>% 
  relocate(translocation_id, .after = site_id)
```

### Model specification

```{r}
m1d_allna <- stan_glmer(
  surv ~ length_z + snowt_z + snowt1_z + day_z + elev_z + first_c + male_c + donor + (1 | translocation_id),
  data = d1_allna,
  family = binomial,
  chains = 4, 
  iter = 5000*2, 
  cores = 4,
  seed = 84735)
```

### List model variable names

```{r}
get_variables(m1d_allna)
```

### Priors & chain diagnostics

```{r}
prior_summary(m1d_allna)

b1_allna <- c("(Intercept)", "length_z", "snowt_z", "snowt1_z", "day_z", "elev_z", "first_c0.349", "male_c0.549", "donor70567", "donor72996")

mcmc_trace(m1d_allna, size = 0.1, pars = b1_allna)
mcmc_dens_overlay(m1d_allna, pars = b1_allna)

summary(m1d_allna)
```

### Posterior predictive check

```{r}
pp_check(m1d_allna, plotfun = "stat") +
    xlab("Frog survival rate")
```

### Posterior analysis

```{r}
tidy(m1d_allna, conf.int = TRUE, conf.level = 0.90)

mcmc_areas(m1d_allna, area_method = "equal height", prob = 0.90, pars = b1_allna) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors & translocation_id as grouping variable",
          "Posterior distributions with medians & 90% uncertainty intervals")

mean(bayes_R2(m1d_allna))

# Frog survival by translocation_id
mcmc_areas_ridges(m1d_allna, regex_pars = "translocation_id") +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Frog survival for each translocation_id")
```

-   Inclusion of all records (i.e., 12 records where bd_load = NA that were excluded in primary analysis) did not change the results.

## ANALYSIS 4: Calculate parameters for viability analysis

### Estimate subadult survival

#### Connect to database

-   Users without credentials to access database can start with `subadult_survival.csv` saved below.

```{r}
con = dbConnect(dbDriver("PostgreSQL"),
                host = Sys.getenv("host"), 
                port = Sys.getenv("port"), 
                dbname = Sys.getenv("dbname"),
                user = rstudioapi::askForPassword("user name"),
                password = rstudioapi::askForPassword("password")
               )
```

#### Retrieve Bd data for translocated frogs from database

-   Used 70550 because of its long and relatively intensive survey history.

```{sql, connection=con, output.var="surv"}
select
  site_id,
  visit_date,
  species,
  visual_life_stage,
  visual_animal_state,
  count
from visit
inner join survey on visit_id = visit.id
inner join visual_survey on survey_id = survey.id  
where site_id in (70550) and
  species = 'ramu' and
  visual_life_stage = 'subadult' and
  visual_animal_state = 'alive'
order by site_id, visit_date
```

#### Disconnect from database

```{r}
dbDisconnect(con) 
rm("con")
```

#### Save raw dataset

```{r}
surv %>% write_csv(here::here("data", "raw", "subadult_survival.csv"))
```

#### Calculate maximum annual counts

```{r}
surv <- surv %>% 
  group_by(visit_date) %>% 
  summarize(count_sum = sum(count)) %>% 
  ungroup() %>% 
  mutate(year = year(visit_date)) %>% 
  group_by(year) %>% 
  summarize(count_max = max(count_sum))
```

#### Extract number of new recruits from mrmr model

```{r}
model <- read_rds(here::here("data", "clean", "70550_model.rds"))

get_numeric_indices <- function(string) {
  idx_mat <- stringr::str_extract_all(string, "[0-9]+", simplify = TRUE)
  colnames(idx_mat) <- paste0("index_", seq_len(ncol(idx_mat)))
  tibble::as_tibble(idx_mat) %>%
    dplyr::mutate_all(readr::parse_integer)
}

survey_prim_periods <- model$data$surveys %>%
  group_by(.data$primary_period) %>%
  filter(.data$secondary_period == min(.data$secondary_period)) %>% 
  ungroup()

p <- model$m_fit$draws("B", format = "draws_df") %>%
  tidyr::pivot_longer(tidyselect::starts_with("B")) %>%
  suppressWarnings() %>%
  mutate(
    get_numeric_indices(.data$name),
    primary_period = .data$index_1 + 1
  ) %>%
  group_by(.data$primary_period) %>%
  summarize(lo = quantile(.data$value, .025),
            med = median(.data$value),
            hi = quantile(.data$value, .975),
            .groups = "drop") %>%
  left_join(survey_prim_periods, by = "primary_period")

recruit <- p %>% 
  group_by(year) %>% 
  summarize(recruit_sum = sum(med))
```

#### Create final dataset

```{r}
subadult_survival <- surv %>% 
  inner_join(recruit, by = c("year"))
```

#### Calculate subadult survival

-   Assumes that p(detection) for subadults is 0.25.

```{r}
subadult_sum <- subadult_survival %>% 
  filter(between(year, 2008, 2020)) %>% 
  summarize(sum(count_max))

recruit_sum <- subadult_survival %>% 
  filter(between(year, 2009, 2021)) %>% 
  summarize(sum(recruit_sum))

survival_sub <- recruit_sum / (subadult_sum * 4) 
```

### Estimate 1st-year tadpole survival

-   Used 74976 due to its shallow habitats that allow accurate tadpole counts.

-   1st-year tadpole counts are from field notebooks.

```{r}
surv_tad <- tibble(site_id = 74976, year = 2019:2021, count = c(250, 300, 305))
```

#### Extract adult abundance from mrmr model

-   Used minimum survey date per year to estimate number of females during breeding season.

```{r}
model <- read_rds(here::here("data", "clean", "74976_model.rds"))

get_numeric_indices <- function(string) {
  idx_mat <- stringr::str_extract_all(string, "[0-9]+", simplify = TRUE)
  colnames(idx_mat) <- paste0("index_", seq_len(ncol(idx_mat)))
  tibble::as_tibble(idx_mat) %>%
    dplyr::mutate_all(readr::parse_integer)
}

survey_prim_periods <- model$data$surveys %>%
  group_by(.data$primary_period) %>%
  filter(.data$secondary_period == min(.data$secondary_period)) %>% 
  ungroup()

p <- model$m_fit$draws("N", format = "draws_df") %>%
        tidyr::pivot_longer(tidyselect::starts_with("N")) %>%
        suppressWarnings() %>%
        mutate(
          get_numeric_indices(.data$name),
          primary_period = .data$index_1 + 1
        ) %>%
        group_by(.data$primary_period) %>%
        summarize(lo = quantile(.data$value, .025),
                  med = median(.data$value),
                  hi = quantile(.data$value, .975),
                  .groups = "drop") %>%
        left_join(survey_prim_periods, by = "primary_period")

abund_ad <- p %>% 
  group_by(year) %>% 
  summarize(survey_date = min(survey_date)) %>%
  select(survey_date) %>% 
  inner_join(p, by = c("survey_date")) %>% 
  select(year, med)
```

#### Create final dataset

```{r}
surv_tad <- surv_tad %>% 
  inner_join(abund_ad, by = "year") %>% 
  rename(adults = med)
```

#### Calculate tadpole survival

Assumptions:

-   p(detection) for 1st-year tadpoles is 0.5.

-   \# females is half of total adult abundance

-   female fecundity = 100 eggs

```{r}
surv_tad <- surv_tad %>% 
  mutate(surv_1styr = (count * 2) / ((adults / 2) * 100)) %>% 
  summarize(mean(surv_1styr))
```
