---
title: Reintroduction of disease-resistant individuals allows recovery of an endangered frog in the presence of a lethal fungal disease
author:
  - name: Roland A. Knapp
    affiliation: a,b,1
  - name: Thomas C. Smith
    affiliation: a,b
  - name: Mark Q. Wilber
    affiliation: c
  - name: Andrew P. Rothstein
    affiliation: d,e,2
  - name: Allison Q. Byrne
    affiliation: d,e,2
  - name: Maxwell B. Joseph
    affiliation: f
address:
    
  - code: a
    address: Sierra Nevada Aquatic Research Laboratory, University of California, Mammoth Lakes, CA 93546
  - code: b
    address: Earth Research Institute, University of California-Santa Barbara, Santa Barbara, CA 93106
  - code: c
    address: Department of Forestry, Wildlife, and Fisheries, University of Tennessee, Knoxville, TN 37996
  - code: d
    address: Department of Environmental Science, Policy, and Management, University of California-Berkeley, Berkeley, CA 94720
  - code: e
    address: Museum of Vertebrate Zoology, University of California-Berkeley, Berkeley, CA 94720
  - code: f
    address: Earth Lab, University of Colorado, Boulder, CO 80303
corresponding_author:
  code: 1
  text: "To whom correspondence should be addressed. E-mail: roland.knapp@ucsb.edu"
lead_author_surname: Knapp
equal_authors:
  code: 2
  text: "A.P.R. and A.Q.B. contributed equally to this work."
author_contributions: |
  R.A.K, T.C.S, M.Q.W, A.P.R., A.Q.B, and M.B.J. designed research; R.A.K, T.C.S, M.Q.W, A.P.R., and A.Q.B performed research; R.A.K, M.Q.W, A.P.R., A.Q.B, and M.B.J. analyzed data; M.B.J. contributed software; and R.A.K, T.C.S, M.Q.W, A.P.R., A.Q.B, and M.B.J. wrote the paper.   
conflict_of_interest: |
  The authors declare no conflicts of interest.
abstract: |
  <!-- <=250 words in a single paragraph--> Pharetra pharetra massa massa ultricies mi quis hendrerit dolor magna eget est lorem ipsum dolor sit amet consectetur adipiscing elit pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas integer eget aliquet nibh praesent tristique magna sit amet purus gravida quis blandit turpis cursus in hac habitasse platea dictumst quisque sagittis purus sit amet volutpat consequat mauris nunc congue nisi vitae suscipit tellus mauris a diam maecenas sed enim ut sem viverra aliquet eget sit amet tellus cras adipiscing enim eu turpis egestas pretium aenean pharetra magna ac placerat vestibulum lectus mauris ultrices eros in cursus turpis massa tincidunt dui ut ornare lectus sit amet est placerat in egestas erat imperdiet sed euismod nisi porta lorem mollis aliquam ut porttitor leo a diam sollicitudin tempor id eu nisl nunc mi ipsum faucibus vitae aliquet nec ullamcorper sit amet risus nullam eget felis eget nunc lobortis mattis aliquam faucibus purus in massa tempor nec feugiat nisl pretium fusce id velit ut tortor pretium viverra suspendisse potenti nullam ac tortor vitae purus faucibus ornare suspendisse sed nisi lacus sed viverra tellus in hac habitasse platea dictumst vestibulum rhoncus est pellentesque elit ullamcorper dignissim cras tincidunt lobortis feugiat vivamus at augue eget arcu dictum varius duis at consectetur lorem donec massa sapien faucibus et molestie ac feugiat sed lectus vestibulum mattis ullamcorper velit sed ullamcorper morbi tincidunt ornare massa eget egestas purus viverra accumsan in nisl nisi scelerisque eu ultrices vitae auctor eu augue ut lectus arcu bibendum at.
significance: |
  <!-- <=120 words --> Enim blandit volutpat maecenas volutpat blandit aliquam etiam erat velit scelerisque in dictum non consectetur a erat nam at lectus urna duis convallis convallis tellus id interdum velit laoreet id donec ultrices tincidunt arcu non sodales neque sodales ut etiam sit amet nisl purus in mollis nunc sed id semper risus in hendrerit gravida rutrum quisque non tellus orci ac auctor augue mauris augue neque gravida in fermentum et sollicitudin ac orci phasellus egestas tellus rutrum tellus pellentesque eu tincidunt tortor aliquam nulla facilisi cras fermentum odio eu feugiat pretium nibh ipsum consequat nisl vel pretium lectus quam id leo in vitae turpis massa sed elementum tempus egestas sed sed risus pretium quam vulputate dignissim suspendisse in est ante in.
acknowledgements: |
  Funding for this research was provided by the National Science Foundation, Yosemite National Park, Yosemite Conservancy, and U.S. Geological Survey NRPP program. Erica Hegeman and A. Lindauer were instrumental in assisting with field work, logistics, and permitting. Many summer field assistants also contributed importantly. 
keywords:
  - population recovery
  - amphibian declines
  - emerging infectious disease
  - reintroduction
  - Batrachochytrium dendrobatis
## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle
bibliography: pnas.bib
csl: pnas.csl
lineno: false
output: rticles::pnas_article
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, results='hide')
```

```{r load-packages}
library(tidyverse)
```
<!-- text formatting & citations -->
The amphibian disease chytridiomycosis, caused by the chytrid fungus
*Batrachochytrium dendrobatidis* ("Bd"), is one of the most destructive
wildlife diseases in recorded history. This recently-emerged fungus
[@ohanlon2018recent] is highly pathogenic to a wide range of amphibian
taxa and, by one estimate, has caused the severe decline or extinction
of at least 500 amphibian species [@scheele2019amphibian], with many
more predicted to be at risk [@rodder2009global]. The pathogenesis of
chytridiomycosis is the result of Bd infection disrupting cutaneous
osmoregulatory function, leading to electrolyte imbalance and death
[@voyles2009pathogenesis]. Bd infection is transferred between hosts via
an aquatic flagellated zoospore stage [@longcore1999batrachochytrium]
pulvinar neque laoreet suspendisse interdum consectetur libero id
faucibus nisl tincidunt eget nullam non nisi est sit amet facilisis
magna etiam tempor orci eu lobortis elementum nibh tellus molestie nunc
non blandit massa enim nec dui nunc mattis enim ut tellus elementum
sagittis vitae et leo duis ut diam quam nulla porttitor massa id neque
aliquam vestibulum morbi blandit cursus risus at ultrices mi tempus
imperdiet nulla malesuada pellentesque elit eget gravida cum sociis
natoque penatibus et magnis dis parturient montes nascetur ridiculus mus
mauris vitae ultricies leo integer malesuada nunc vel risus commodo
viverra maecenas accumsan lacus vel facilisis volutpat est velit egestas
dui id ornare arcu odio ut sem nulla pharetra diam sit amet nisl
suscipit adipiscing bibendum est ultricies integer quis auctor elit sed
vulputate mi sit.

amet mauris commodo quis imperdiet massa tincidunt nunc pulvinar sapien
et ligula ullamcorper malesuada proin libero nunc consequat interdum
varius sit amet mattis vulputate enim nulla aliquet porttitor lacus
luctus accumsan tortor posuere ac ut consequat semper viverra nam libero
justo laoreet sit amet cursus sit amet dictum sit amet justo donec enim
diam vulputate ut pharetra sit amet aliquam id diam maecenas ultricies
mi eget mauris pharetra et ultrices neque ornare aenean euismod
elementum nisi quis eleifend quam adipiscing vitae proin sagittis nisl
rhoncus mattis rhoncus urna neque viverra justo nec ultrices dui sapien
eget mi proin sed libero enim sed faucibus turpis in eu mi bibendum
neque egestas congue quisque egestas diam in arcu cursus euismod quis
viverra. <!-- Inline code & embedded plots -->

# Results {.unnumbered}

```{r inline-values, include=FALSE}
frog_translocation <- read_csv(here::here("data", "clean", "frog_translocation.csv")) %>% 
  mutate(site_id = as.character(site_id)) 
count_translocation <- frog_translocation %>% distinct(site_id, date) %>% count()
count_site <- frog_translocation %>% distinct(site_id) %>% count()
min_elev <- min(frog_translocation$elevation)
max_elev <- max(frog_translocation$elevation)
bdload_groupmed <- 
  frog_translocation %>% group_by(site_id, date) %>% summarize(median_bdload = median(bd_load, na.rm = TRUE)) 
bd_min <- min(bdload_groupmed$median_bdload) %>% round()
bd_max <- max(bdload_groupmed$median_bdload) %>% round() %>% format(scientific = FALSE)
```

```{r lw-plot, include=FALSE, dev="png", fig.width=3.3, fig.height=3.3, dpi=300, fig.show="hide"}
lw_plot <- frog_translocation %>% ggplot(aes(x = length, y = weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Length (mm)", 
       y = "Weight (g)") +
  theme_classic()       
plot(lw_plot)

```

During the 2006--2020 period covered by this study, we conducted
`r count_translocation$n` translocations to `r count_site$n` sites.
Sites ranged in elevation from `r min(frog_translocation$elevation)` to
`r max(frog_translocation$elevation)` m, and median Bd loads of
translocated cohorts ranged from `r bd_min` to `r bd_max` ITS copies per
skin swab. pulvinar neque laoreet suspendisse interdum consectetur
libero id faucibus nisl tincidunt eget nullam non nisi est sit amet
facilisis magna etiam tempor orci eu lobortis elementum nibh tellus
molestie nunc non blandit massa enim nec dui nunc mattis enim ut tellus
elementum sagittis vitae et leo duis ut diam quam nulla porttitor massa
id neque aliquam vestibulum morbi blandit cursus risus at ultrices mi
tempus imperdiet nulla malesuada pellentesque elit eget gravida cum
sociis natoque penatibus et magnis dis parturient montes nascetur
ridiculus mus mauris vitae ultricies leo integer malesuada nunc vel
risus commodo viverra maecenas accumsan lacus vel facilisis volutpat est
velit egestas dui id ornare arcu odio ut sem nulla pharetra diam sit
amet nisl suscipit adipiscing bibendum est ultricies integer quis auctor
elit sed vulputate mi sit.

amet mauris commodo quis imperdiet massa tincidunt nunc pulvinar sapien
et ligula ullamcorper malesuada proin libero nunc consequat interdum
varius sit amet mattis vulputate enim nulla aliquet porttitor lacus
luctus accumsan tortor posuere ac ut consequat semper viverra nam libero
justo laoreet sit amet cursus sit amet dictum sit amet justo donec enim
diam vulputate ut pharetra sit amet aliquam id diam maecenas ultricies
mi eget mauris pharetra et ultrices neque ornare aenean euismod
elementum nisi quis eleifend quam adipiscing vitae proin sagittis nisl
rhoncus mattis rhoncus urna neque viverra justo nec ultrices dui sapien
eget mi proin sed libero enim sed faucibus turpis in eu mi bibendum
neque egestas congue quisque egestas diam in arcu cursus euismod quis
viverra. Following log~10~-transformation, frog weight was linearly
related to frog length across most of the range of frog lengths.
However, for lengths \> 56 mm nearly all residuals from the
length:weight regression line were mostly positive.

![](`r knitr::fig_chunk('lw-plot', 'png')`)

Figure 1. For translocated mountain yellow-legged frogs, the relationship between length and weight (on a log~10~-log~10~
scale).

## Evolution of Reduced Disease Susceptibility {#results-subheading1 .unnumbered}

## Reintroduction of Disease-resistant Frogs {#results-subheading2 .unnumbered}

## Viability of Reestablished Populations {#results-subheading3 .unnumbered}

## Conceptual Model of Population Establishment {#results-subheading4 .unnumbered}

# Discussion {.unnumbered}

## Conclusions {#discussion-subheading1 .unnumbered}

# Materials and Methods {#matmethods .unnumbered}

amet mauris commodo quis imperdiet massa tincidunt nunc pulvinar sapien
et ligula ullamcorper malesuada proin libero nunc consequat interdum
varius sit amet mattis vulputate enim nulla aliquet porttitor lacus
luctus accumsan tortor posuere ac ut consequat semper viverra nam libero
justo laoreet sit amet cursus sit amet dictum sit amet justo donec enim
diam vulputate ut pharetra sit amet aliquam id diam maecenas ultricies
mi eget mauris pharetra et ultrices neque ornare aenean euismod
elementum nisi quis eleifend quam adipiscing vitae proin sagittis nisl
rhoncus mattis rhoncus urna neque viverra justo nec ultrices dui sapien
eget mi proin sed libero enim sed faucibus turpis in eu mi bibendum
neque egestas congue quisque egestas diam in arcu cursus euismod quis
viverra. Following log\~10-transformation, frog weight was linearly
related to frog length across most of the range of frog lengths.
However, for lengths \> 56 mm nearly all residuals from the
length:weight regression line were mostly. positive.
<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->

```{=tex}
\showmatmethods
\showacknow
\pnasbreak
```
\bigskip