---
format: docx
editor: visual
csl: science.csl
bibliography: translocation.bib
---

## Title: Reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### Authors:

Roland A. Knapp^1,2^\*, Allison Q. Byrne^3,4^, Mark Q. Wilber^5^, Maxwell B. Joseph^6,7^, Thomas C. Smith^1,2^, Andrew P. Rothstein^3,4^, Rob L. Grasso^8^, Erica Bree Rosenblum^3,4^ (exact order to be determined)

### Affiliations:

^1^Sierra Nevada Aquatic Research Laboratory, University of California; Mammoth Lakes, CA 93546, USA

^2^Earth Research Institute, University of California; Santa Barbara, CA, 93106, USA

^3^Department of Environmental Science, Policy, and Management, University of California; Berkeley, CA 94720, USA

^4^Museum of Vertebrate Zoology, University of California; Berkeley, CA 94720, USA

^5^Department of Forestry, Wildlife, and Fisheries, University of Tennessee Institute of Agriculture; Knoxville, TN 37996, USA

^6^Earth Lab, University of Colorado; Boulder, CO 80303, USA

^7^NCX; San Francisco, CA, XXXXX, USA

^8^Resources Management and Science, Yosemite National Park; El Portal, CA 95318, USA

\*Corresponding author: roland.knapp\@ucsb.edu

### Abstract:

### One Sentence Summary

### Main Text:

Severe effects of disease on global biodiversity (bats, primates, amphibians, corals, sea stars, etc.)

Host evolution and population persistence/recovery

-   Species recovery in the face of disease will depend on the ability of a species to move or change.

-   Rapid evolution termed "evolutionary rescue" can reverse ongoing declines in some species (Golas et al. 2021; Searle et al. 2021).
    (Not sure if Allie wants to go the ER direction)

-   ER has been proposed as a mechanism of recovery in wildlife populations decimated by disease (Epstein et al. 2016, Schiebelhut et al. 2018).

    -   Tasmanian devils: Epstein et al. (2016), Wells et al., Ecology (2019)

    -   Frogs in Australia: Grogan et al. (2018), McKnight et al. (2020), Newell et al. (2013), Scheele et al., Biological Conservation (2014)

    -   Frogs in Panama: Voyles et al. (2018)

    -   Frogs in the Sierra Nevada: Knapp et al. (2016)

    -   Sea Star Wasting Disease: Menge et al., Plos One (2016), Schiebelhut et al.
        PNAS (2018)

-   **Potential knowledge gap**: Subsequent recoveries from disease-induced declines following observed genetic changes have rarely \[never?\]
    been documented in wildlife populations (Sea stars: Moritsch and Raimondi 2018?).

Population reestablishment using reintroduction of resistant hosts

-   "Rescued" populations provide a source to repatriate extirpated populations and disentangle the effects of the local environment and genetics on host recoveries.

-   Mendelson et al. (2019)

-   Scheele et al. (2021)

-   Joseph and Knapp (2018)

-   **Potential knowledge gap/opportunity/selling point**: For rescued populations to contribute to landscape-scale species recovery, the fitness benefits gained through selection should manifest beyond the natal environment, allowing for the long-term growth and persistence of incipient populations in novel environmental conditions.
    A unique attribute of this study is that by combining repeated translocations of "rescued" populations, longitudinal surveys, mark-recapture analysis, and population modeling we demonstrate the capacity of "rescued" populations to persist across a landscape of disease.
    Thus, ER combined with translocations promotes the recovery of a species on the brink of disease-induced extinction.

[This is excellent. Thanks for adding it.]: #

### References and Notes:

### Acknowledgements:

\#### Funding

\#### Author contributions:

\#### Competing interests:

\#### Data and materials availability

### Supplementary Materials:

### Fig. #.

### Table #.

### Notes (author use only - not part of manuscript)

Science:

-   Research articles

    -   Present a major advance.

    -   Up to \~4500 words, including references, notes and captions.
        Corresponds to approximately 5 printed pages in journal.

    -   Up to 6 figures or tables, and about 40 references.

-   Reports

    -   New research results of broad significance.

    -   Up to \~2500 words, including references, notes and captions.
        Corresponds to approximately 3 printed pages in journal.

    -   Up to 4 figures or tables, and about 30 references.

Nature

-   Articles

    -   Original reports whose conclusions represent a substantial advance in understanding of an important problem and have immediate, far-reaching implications.

    -   Biological sciences papers do not normally exceed 8 pages.
        This would be approximately 4300 words and 5-6 modest display items.

PNAS

-   Research reports

    -   Original research of exceptional importance.

    -   Preferred length is 6 pages, but up to 12 pages is allowed.
        6 page article is approximately 4,000 words, 50 references, and 4 medium-size figures or tables.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code.
You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
