---
format: docx
editor: visual
csl: nature.csl
bibliography: translocation.bib
---

## Title: Reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

<!--# 75 characters. Suggestions to make this more compelling? Use "Evolutionary rescue"? -->

### Authors:

Roland A. Knapp^1,2^\*, Allison Q. Byrne^3,4^, Mark Q. Wilber^5^, Maxwell B. Joseph^6,7^, Thomas C. Smith^1,2^, Andrew P. Rothstein^3,4^, Rob L. Grasso^8^, Erica Bree Rosenblum^3,4^

<!--# Exact author order to be determined) -->

### Affiliations:

^1^Sierra Nevada Aquatic Research Laboratory, University of California, Mammoth Lakes, CA, USA

^2^Earth Research Institute, University of California, Santa Barbara, CA, USA

^3^Department of Environmental Science, Policy, and Management, University of California. Berkeley, Berkeley, CA, USA

^4^Museum of Vertebrate Zoology, University of California, Berkeley, Berkeley, CA, USA

^5^Department of Forestry, Wildlife, and Fisheries, University of Tennessee Institute of Agriculture, Knoxville, TN, USA

^6^Earth Lab, University of Colorado, Boulder, CO, USA

^7^NCX, San Francisco, CA, USA

^8^Resources Management and Science, Yosemite National Park, El Portal, CA , USA

\*Corresponding author: roland.knapp\@ucsb.edu

### Abstract

```{=html}
<!--# No more than 200 words, follow structure described in instructions for authors. 2-3 sentences of basic-level introduction to the field; a brief account of the background and rationale of the work; a statement of the main conclusions (introduced by the phrase 'Here we show' or its equivalent);
 and finally, 2-3 sentences putting the main findings into general context so it is clear how the results described in the paper have moved the field forwards. See [annotated example](https://www.nature.com/documents/nature-summary-paragraph.pdf).  -->
```
### Main

<!--# subheadings can be used, can be "teasers" of results (up to 40 characters) -->

Severe effects of disease on global biodiversity (bats, primates, amphibians, corals, sea stars, etc.)

Host evolution and population persistence/recovery

-   Species recovery in the face of disease will depend on the ability of a species to move or change.

-   Rapid evolution termed "evolutionary rescue" can reverse ongoing declines in some species (Golas et al. 2021; Searle et al. 2021). (Not sure if Allie wants to go the ER direction)

-   ER has been proposed as a mechanism of recovery in wildlife populations decimated by disease (Epstein et al. 2016, Schiebelhut et al. 2018).

    -   Tasmanian devils: Epstein et al. (2016), Wells et al., Ecology (2019)

    -   Frogs in Australia: Grogan et al. (2018), McKnight et al. (2020), Newell et al. (2013), Scheele et al., (2014)

    -   Frogs in Panama: Voyles et al. (2018)

    -   Frogs in the Sierra Nevada: Knapp et al. (2016) @knapp2016

    -   Sea Star Wasting Disease: Menge et al., Plos One (2016), Schiebelhut et al. PNAS (2018)

    -   White-nose syndrome: Gignoux-Wolfsohn et al. (2021) Molecular ecology.

-   **Potential knowledge gap**: Subsequent recoveries from disease-induced declines following observed genetic changes have rarely \[never?\] been documented in wildlife populations (Sea stars: Moritsch and Raimondi 2018?).

Population reestablishment using reintroduction of resistant hosts

-   "Rescued" populations provide a source to repatriate extirpated populations and disentangle the effects of the local environment and genetics on host recoveries.

-   Mendelson et al. (2019)

-   Scheele et al. (2021)

-   Joseph and Knapp (2018)

-   **Potential knowledge gap/opportunity/selling point**: For rescued populations to contribute to landscape-scale species recovery, the fitness benefits gained through selection should manifest beyond the natal environment, allowing for the long-term growth and persistence of incipient populations in novel environmental conditions. A unique attribute of this study is that we describe the genetic changes associated with evolutionary rescue, and combine repeated translocations of "rescued" populations, longitudinal surveys, mark-recapture analysis, and population modeling to demonstrate the capacity of "rescued" populations to persist across a landscape of disease. Thus, ER combined with translocations promotes the recovery of a species on the brink of disease-induced extinction. These results may have broad implications for the many species worldwide that are threatened with extinction by the emergence of novel diseases.

Description of MYL frog species complex, impacts of Bd, and evidence of recent recovery.

#### Frog evolution in response to Bd

#### Frog translocations

#### Population viability modeling

Given observed yearly adult survival probabilities, our stage-structured model predicted that six of our twelve translocated population should experience a long-run growth rate $\lambda$ near or greater than 1 in the presence of Bd (Fig. XA; Fig. SXA; median predicted $\lambda$ ranges from 0.95-1.16 for these six populations). When the probability of successful recruitment $\omega$ was greater than 0.1 (the conservative estimate we used as a baseline value in our model), long-run growth rate of these populations could be notably greater than 1 (Fig. XA). In contrast, for the other six translocated populations where yearly adult survival was less than 0.5, our modeled predicted that it was unlikely that long-run growth rate in these populations was greater than 1, even when the probability of successful recruitment to the adult lifestage was near 0.4 (Fig. XA). A sensitivity analysis of long-run growth rate $\lambda$ to recruitment probability $\omega$ and yearly adult survival probability of naturally recruited individuals $\sigma_{A_R}$ showed that $\lambda$ was most sensitive to small changes in $\sigma_{A_R}$ (Fig. SXA).

Given demographic stochasticity (i.e., probabilistic births and deaths of individuals in a population) and environmental variability in the probability of adult recruitment $\omega$, our model predicted that, following a a single introduction of 40 adult individuals into a population, the six populations with the highest adult survival probabilities had 50 year extinction probabilities of less 1 (Fig. XB) -- in other words, even under conservative assumptions about recruitment probability $\omega$, these populations had some probability of persistence over 50 years following a single introduction. For these six populations, as $\omega$ increased to 0.3, 50 year extinction probability was less 0.5, indicating strong potential for long-term persistence in the presence of Bd and environmental variability in recruitment success (Fig. XB). In contrast, for the six populations where $\sigma_{A_R} < 0.5$, extinction probability over 50 years was always predicted to be 1. To test the validity of our model predictions, we demonstrated that our stochastic model could describe the general recovery trajectory of our translocated population with the longest surveillance history (Fig. SXB; population 70550, surveyed for 16 years).

### Figures and figure legends

<!--# In initial submission, include figure legends with their corresponding figures. -->

<!--# Discuss referencing and in-text generation of figures.  We don't want to re-run the population viability modeling figures each time we compile the manuscript.  A makefile could be a good solution. -->

![Figure X: A. The predicted long-run growth rate $\lambda$ for different values of yearly adult survival probability $\sigma_{A_R}$ and probability or recruitment $\omega$, given the parameterized, deterministic model for translocated $R. sierrae$. Color points show the predicted $\lambda$ values for the twelve translocated populations when probability of recruitment to adults is $\omega = 0.1$ (given by the dashed line). B. The predicted fifty-year extinction probabilities of the twelve translocated populations given demographic stochasticity, environmental variability in $\omega$ and different mean values of $\omega$. The dashed line shows where $\omega = 0.1$, for comparison to A.](figures/pop_viability_figures_for_manuscript.jpg)

### References

::: {#refs}
:::

### Methods

<!--# Typically less than 3000 words -->

#### Frog evolution in response to Bd

<!--# subheadings can be used, up to 40 characters -->

We used exome capture methods to compare genomes of frogs in populations with different histories of Bd exposure, in an effort to putative genes under selection that are linked to reduced frog susceptibility to Bd-caused lethal and sub-lethal effects. Specifically, we compared genomes of frogs that have never experienced a Bd-caused epizootic ("Bd-naive") versus populations that experienced a Bd epizootic during the past several decades and have since recovered to varying degrees ("recovering). These populations persist in an enzootic state, characterized by high Bd prevalence and Bd infection intensities (load) in adults that are well below the level expected to cause frog mortality. Bd-exposure histories of these populations are based on a decade or more of data from VES and Bd surveillance using skin swabbing.

<!--# Mention sampled declining populations also? -->

DNA samples were collected via buccal swabbing of xx-xx frogs from seven Bd-naive populations <!--# Dusy 10220 not included due to small sample size? --> and seven recovering populations. These populations represented a wide latitudinal range in the southern Sierra Nevada from northern Yosemite National Park to south of Sequoia National Park (Figure Sx), and the latitudinal range of the two population types was similar.

<!--# Create map of sample sites -->

#### Frog translocations

Translocations of *R. sierrae* were conducted in Yosemite National Park (California, USA) as part of an effort to reestablish populations of this endangered species. We identified donor populations from which adult frogs could be collected using several years of visual encounter surveys (VES) and skin swab collections (Knapp et al. 2016). Results indicated that these populations contained several hundred *R. sierrae* adults and hundreds or thousands of tadpoles, and had recovered from previous Bd-caused declines. Bd prevalence donor populations is high and Bd infection intensity (load) is below the level at which frog mortality is predicted (calculate mean prevalence and load). Recipient sites to which frogs were translocated were chosen based on previous *R. sierrae* presence (determined from VES and/or museum records) or characteristics that suggested high quality habitat for this species (Knapp 2005).

Between 2006 and 2020 we conducted 1-4 translocations of adult frogs to 13 recipient sites (Table S1). In each translocation, adult frogs ($\geq$ 40 mm snout-vent length) were collected from one of three donor populations and measured, weighed, swabbed, and PIT tagged. Frogs were transported to the recipient site either on foot or via helicopter. Following release, translocated populations were visited one to six times per summer (calculate min, max, mean) and assessed using diurnal capture-mark-recapture (CMR) surveys and VES. CMR surveys allowed quantification of adult frog survival, adult population size, and recruitment of new adults, and VES provided estimates of tadpole and subadult abundance.

<!--# Create table of translocation sites, years, frog numbers. Could also create a map if necessary -->

During 2006-2012, CMR surveys were conducted conducted on a single day (primary period) per site visit, during which all habitats were searched repeatedly for adult frogs. Frog were captured using handheld nets, identified via their PIT tag (or tagged if they were untagged), measured, weighed, swabbed, and released at the capture location. During 2013-2022, we used a robust design in which all habitats were searched during several consecutive days (secondary periods - calculate min, max, mean), and frogs were processed as described above. Within a primary period, frogs that were captured during more than one secondary period were measured, weighed, and swabbed only during the first capture, and identified only during subsequent captures. Frog skin swabs were analyzed using standard Bd DNA extraction and qPCR methods as described in Joseph and Knapp (2018).

During each site visit, we conducted VES either immediately before CMR surveys or during the first day of CMR surveys. VES was conducted by walking the entire water body perimeter, first 100 m of each inlet and outlet stream, and any fringing ponds and wetlands, and counting all *R. sierrae* tadpoles and subadults. These *R. sierrae* life stages have high detectablility, and counts are highly repeatable and provide estimates of relative abundance.

#### Population viability modeling

**Model description**

To determine the implications of acquired Bd resistance on long-term population viability, we developed a dynamic, population model for *R. sierrae*. Our central question was: how does the magnitude and variation in observed adult survival probability across translocated populations affect the long-term persistence probability of populations? We developed a model that tracked seven state variables of a *R. sierrae* population: the density of translocated adults ($A_T$), the density of adults naturally recruited into the population ($A_R$), the density of first year larvae ($L_1$), the density of second year larvae ($L_2$), the density of third year larvae ($L_3$), the density of first year juveniles ($J_1$), and the density of second year juveniles ($J_2$). We divided adults into two classes $A_T$ and $A_R$ because there was evidence that the survival probability of translocated adults and naturally recruited adults differed.

We modeled the dynamics of these seven state variables using a discrete-time, stage-structured model where a time step is one year. The dynamics are given by

$$
\begin{bmatrix}
L_1 \\
L_2 \\
L_3 \\
J_1 \\
J_2 \\
A_R \\ 
A_T
\end{bmatrix}(t + 1) = 
\begin{bmatrix}
  0 & 0 & 0 & 0 & 0 & \sigma_{A_R} p_F F & \sigma_{A_T} p_F F \\
  \sigma_{L1} p_{L1} & 0 & 0 & 0 & 0 & 0 & 0 \\
  0 & \sigma_{L2} p_{L2} & 0 & 0 & 0 & 0 & 0\\
  \sigma_{L1} (1 - p_{L1}) & \sigma_{L2} (1 - p_{L2}) & \sigma_{L3} & 0 & 0 & 0 & 0 \\
  0 & 0 & 0 & \sigma_{J1} p_{J1} & 0 & 0 & 0 \\
  0 & 0 & 0 & \sigma_{J1} (1 - p_{J1}) \omega & \sigma_{J2} \omega & \sigma_{A_R} & 0 \\
  0 & 0 & 0 & 0 & 0 & 0 & \sigma_{A_T} 
\end{bmatrix} \begin{bmatrix}
L_1 \\
L_2 \\
L_3 \\
J_1 \\
J_2 \\
A_R \\
A_T \\
\end{bmatrix}(t) 
$$

The parameters in this model are yearly survival probability $\sigma_{*}$, where the "\*" indicates a particular state variable, the probability that a female frog reproduces in a given year $p_F$, the number of eggs produced by a female frog in a year that successfully hatch $F$, the probability of a first year tadpole remaining as a tadpole $p_{L1}$, the probability of a second year tadpole remaining as a tadpole $p_{L2}$, and the probability of a first year juvenile remaining as a juvenile $p_{J1}$. The parameter $\omega$ gives the probability of a juvenile successfully recruiting to an adult.

In this model we ignore density-dependent recruitment because we were interested in the growth of the population from an initial reintroduction and whether this growth was sufficient to prevent extinction over 50 years following the introduction. We also did not directly consider the dynamics of Bd in this model. We made this decision because i) all translocated populations are infected with Bd at high prevalence ii) host density does not seem to play a significant role in multi-year Bd infection dynamics in this system. Thus, ignoring Bd infection dynamics and instead assuming all host vital rates are in the presence of high Bd prevalence significantly simplifies the model without much loss of realism.

**Model analysis**

**Incorporating yearly variability in vital rates**

We extracted yearly survival probabilities for translocated adults $\sigma_{A_T}$ and naturally recruited adults $\sigma_{A_R}$ from the mark-recapture model described elsewhere. While we observed yearly variability in adult survival *within* a population, the magnitude of this variability was small compared to *among* population variability. Thus, we did not include yearly within-population variability in adult survival in this analysis.

However, within a population there was substantial yearly variability in the successful recruitment of adults, greater than what we would expect from Poisson variability about a mean value. Therefore, we allowed for yearly variability in the probability of successful recruitment $\omega$ (more details below).

**Estimating model parameters**

The baseline parameter values for the model and how they were estimated are given in Table 1. Parameters $\sigma_{A_T}$ and $\sigma_{A_R}$ were extracted directly from our mark-recapture model described elsewhere. For populations where we had a sufficient number of pit-tagged, naturally recruited adults, we observed that $\sigma_{A_T}$ and $\sigma_{A_R}$ could be notably different, with $\sigma_{A_R} > \sigma_{A_T}$. For populations where we had not pit-tagged any naturally recruited adults, we could not directly estimate $\sigma_{A_R}$ and set $\sigma_{A_R} = \sigma_{A_T}$.

**Model analysis and simulation**

We performed four analyses on our model. First, we considered a deterministic version of our model with no yearly heterogeneity in probability of successful recruitment $\omega$ and calculated the predicted long-run growth rate $\lambda$ of a population for different values of $\sigma_{A_R}$ and $\omega$. We then fixed $\omega = 0.1$ and calculated the predicted growth rate of our six populations. Second, we performed a local sensitivity analysis on $\lambda$ with respect to parameters $\omega$, $\sigma_{A_R}$, and $F$ to determine how small changes in these parameters could influence the long-run deterministic growth rate of populations.

Third, we defined version of the model with demographic and environmental stochasticity, where environmental stochasticity was represented by among year variability in $\omega$. We used this model to simulate a one time introduction of 40 translocated adult frogs. We ran this simulation 1000 times for each population and computed the probability of a population becoming extinct after 50 years given the observed parameter values and environmental stochasticity in $\omega$. We varied the mean recruitment probability $\omega$ from 0.05 and 0.3 and drew values of $\omega$ each year from a beta distribution with a dispersion parameter of $\phi = 2$ (when $\omega = 0.5$ and $\phi = 2$ the beta distribution is uniform between 0 and 1).

Finally, we assessed whether our stochastic model could reproduce observed trajectories of population recovery. We focused on population 70550 because this was our longest mark-recapture time series for a translocated population and because this population shows evidence of recovery. We simulated our model for 16 years, repeating the simulation 50,000 times. For each run and each year, we drew $\omega$ from a uniform distribution between 0 and 1 (or a beta distribution with mean 0.5 and $\phi = 2$). Using Approximate Bayesian Computing and rejection sampling (Kosmala et al. 2015, Ecology), we identified the top 2% of trajectories (i.e., 1000 trajectories) that minimized the sum of squared errors between the observed and predicted data. The $\omega$ values associated with these "best" trajectories represented an approximate posterior distribution (Beaumont 2010). Using these best fit trajectories, we assessed whether our model could qualitatively describe the patterns of recovery in the observed data for population 70550.

### Data availability

### References

### Acknowledgements

### Author contributions

### Competing interests

The authors declare no competing interests.

### Additional information

Supplementary Information is available for this paper.

Correspondence and requests for materials should be addressed to Roland Knapp.

Peer review information.

Reprints and permissions information is available at www.nature.com/reprints.

### Supplementary Information

### Notes (author use only - not part of manuscript)

Science (impact factor = 63.714)

-   Research articles

    -   Present a major advance.

    -   Up to approximately 4500 words, including references, notes and captions. Corresponds to approximately 5 printed pages in journal (approxiumately 11 manuscript pages).

    -   Up to 6 figures or tables, and about 40 references.

-   Reports

    -   New research results of broad significance.

    -   Up to \~2500 words, including references, notes and captions. Corresponds to approximately 3 printed pages in journal (approximately 6 manuscript pages).

    -   Up to 4 figures or tables, and about 30 references.

Nature (impact factor = 69.8)

-   Articles

    -   Original reports whose conclusions represent a substantial advance in understanding of an important problem and have immediate, far-reaching implications.

    -   Biological sciences papers do not normally exceed 8 pages. This would be approximately 4300 words and 5-6 modest display items.

    -   This [article](https://www.nature.com/articles/s41586-022-04824-9) provides a potentially useful (e.g., recent) example of formatting and structure for a paper focused on ecology/evolution

PNAS (impact factor = 12.77)

-   Research reports

    -   Original research of exceptional importance.

    -   Preferred length is 6 pages, but up to 12 pages is allowed. 6 page article is approximately 4,000 words, 50 references, and 4 medium-size figures or tables.

Nature Communications (impact factor = 17.694)

-   Article

    -   Papers published by the journal represent important advances of significance to specialists within each field. Open access.

    -   The main text (not including abstract, Methods, References and figure legends) is limited to 5,000 words. The maximum title length is 15 words.

Science Advances (impact factor = 14.957)

-   Publishes impactful research papers and reviews in any area of science. Journal extends the capacity of *Science* magazine to identify and promote significant advances in science and engineering. Open access.

-   Research article

    -   Should present a major advance.

    -   Articles can be short of long format, up to 15,000 words, with up to 10 figures and/or tables.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
