---
format: docx
editor: visual
csl: science.csl
bibliography: translocation.bib
reference-doc: translocation_style_ref.docx
---

## Title: Reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### Authors:

Roland A. Knapp^1,2^\*, Allison Q. Byrne^3,4^, Mark Q. Wilber^5^, Maxwell B. Joseph^6,7^, Thomas C. Smith^1,2^, Andrew P. Rothstein^3,4^, Rob L. Grasso^8^, Erica Bree Rosenblum^3,4^ (exact order to be determined)

### Affiliations:

^1^Sierra Nevada Aquatic Research Laboratory, University of California; Mammoth Lakes, CA 93546, USA

^2^Earth Research Institute, University of California; Santa Barbara, CA, 93106, USA

^3^Department of Environmental Science, Policy, and Management, University of California; Berkeley, CA 94720, USA

^4^Museum of Vertebrate Zoology, University of California; Berkeley, CA 94720, USA

^5^Department of Forestry, Wildlife, and Fisheries, University of Tennessee Institute of Agriculture; Knoxville, TN 37996, USA

^6^Earth Lab, University of Colorado; Boulder, CO 80303, USA

^7^NCX; San Francisco, CA, XXXXX, USA

^8^Resources Management and Science, Yosemite National Park; El Portal, CA 95318, USA

\*Corresponding author: roland.knapp\@ucsb.edu

### Abstract:

### One Sentence Summary

### Main Text:

Severe effects of disease on global biodiversity (bats, primates, amphibians, corals, etc.)

Host evolution and population persistence/recovery

-   Tasmanian devils: Epstein et al. (2016)

-   Frogs in Australia: Grogan et al. (2018), McKnight et al. (2020)

-   Frogs in Panama: Voyles et al. (2018)

-   Frogs in the Sierra Nevada: Knapp et al. (2016)

Population reestablishment using reintroduction of resistant hosts

-   Mendelson et al. (2019)

-   Scheele et al. (2021)

-   Joseph and Knapp (2018)

### References and Notes:

### Acknowledgements:

\#### Funding

\#### Author contributions:

\#### Competing interests:

\#### Data and materials availability

### Supplementary Materials:

### Fig. #.

### Table #.

### Notes (author use only - not part of manuscript)

Science:

-   Research articles

    -   Present a major advance.

    -   Up to \~4500 words, including references, notes and captions.
        Corresponds to approximately 5 printed pages in journal.

    -   Up to 6 figures or tables, and about 40 references.

-   Reports

    -   New research results of broad significance.

    -   Up to \~2500 words, including references, notes and captions.
        Corresponds to approximately 3 printed pages in journal.

    -   Up to 4 figures or tables, and about 30 references.

Nature

-   Articles

    -   Original reports whose conclusions represent a substantial advance in understanding of an important problem and have immediate, far-reaching implications.

    -   Biological sciences papers do not normally exceed 8 pages.
        This would be approximately 4300 words and 5-6 modest display items.

PNAS

-   Research reports

    -   Original research of exceptional importance.

    -   Preferred length is 6 pages, but up to 12 pages is allowed.
        6 page article is approximately 4,000 words, 50 references, and 4 medium-size figures or tables.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code.
You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
